---
title: "Library Composition Forensics"
keep_md: no
author: "Ryan Ward"
output: html_document
---

Load libraries and constants...

```{r}
library(pacman)

p_load(data.table, viridis, tidyverse, edgeR, scales)
p_load_current_gh("hrbrmstr/hrbrthemes")
annotated_key <- fread("annotated_key.tsv")

```

Determine if the data is good quality by checking for promoter contamination between samples.

```{r, fig.width=10, fig.height=5}

additional <- fread(
	"../additional_sequences/additional.tsv.gz",
	header = FALSE, 
	col.names = c(
		"count",
		"condition",
		"promoter",
		"spacer"))

additional <- additional %>%
	separate(
		condition, 
		into = c("claimed_promoter", "strain"), 
		sep = "_")

forensics.1 <- additional %>% 
	group_by(claimed_promoter, promoter, strain) %>% 
	summarise(count = sum(count)) %>%
	pivot_wider(id_cols = c( "strain", "claimed_promoter"), names_from = promoter, values_from = count) %>%
	unite("cloned_strain", claimed_promoter, strain) 

forensics.1.plot <- forensics.1 %>% 
	select(-cloned_strain) %>% 
	data.matrix %>% t %>% cpm %>% t %>%
	as_tibble %>%
	bind_cols(forensics.1 %>% select(cloned_strain)) %>%
	pivot_longer(!cloned_strain, names_to = "promoter", values_to = "CPM") %>%
	ggplot(aes(fill = promoter, x = cloned_strain, y = CPM)) +
	geom_bar(position = "dodge", stat = "identity", color = "black") +
	theme_ipsum() + 
	scale_fill_viridis(discrete = TRUE, option = "cividis") + 
	ggtitle("Promoters (P1, P2, and P3) found in each strain (CPM)") +
	scale_y_continuous(labels = label_number_si())

print(forensics.1.plot)
```

There is significant contamination of P3 in the sample which contains the P1 library mated into PA14.

Instead of "P1_PA14" use the inoculum from the mouse experiment.

```{r}

additional <- fread(
	"../additional_sequences/additional.tsv.gz",
	header = FALSE, 
	col.names = c(
		"count",
		"condition",
		"promoter",
		"spacer"))

counts.batch1 <- fread(
	"all_counts_seal.tsv.gz",
	header = FALSE,
	col.names = c(
		"promoter",
		"spacer",
		"count",
		"condition"))

```
Remove P1_PA14 counts, replace with Mouse_P1_001 counts, and rename to be more
descriptive: PA_PA14 (round 1). Then separate into two columns
"claimed_promoter" and "strain", which are separated by "_".

```{r, fig.width=10, fig.height=5}

additional <- additional %>% 
	filter(condition != "P1_PA14") %>%
	bind_rows(counts.batch1 %>% filter(condition == "Mouse_P1_003")) %>%
	mutate(
		condition = replace(
			condition, 
			condition == "Mouse_P1_003", 
			"P1_PA14 (round 1)")) %>%
	separate(
		condition, 
		into = c("claimed_promoter", "strain"), 
		sep = "_")

# Perform forensics to see the composition of the strain-promoter combos

forensics.2 <- additional %>% 
	group_by(claimed_promoter, promoter, strain) %>% 
	summarise(count = sum(count)) %>%
	pivot_wider(id_cols = c( "strain", "claimed_promoter"), names_from = promoter, values_from = count) %>%
	unite("cloned_strain", claimed_promoter, strain) 

# Counts per million by each (promoter, claimed_promoter+strain) combo, then 
# magrittr into ggplot and bar plot.

forensics.2.plot <- forensics.2 %>% 
	select(-cloned_strain) %>% 
	data.matrix %>%
	t %>%
	cpm %>%
	t %>%
	as_tibble %>%
	bind_cols(forensics.2 %>% select(cloned_strain)) %>%
	pivot_longer(!cloned_strain, names_to = "promoter", values_to = "CPM") %>%
	ggplot(aes(fill = promoter, x = cloned_strain, y = CPM)) +
	geom_bar(position = "dodge", stat = "identity", color = "black") +
	theme_ipsum() + 
	scale_fill_viridis(discrete = TRUE, option = "cividis") + 
	ggtitle("Promoters (P1, P2, and P3) found in each strain (CPM)") +
	scale_y_continuous(labels = label_number_si())

print(forensics.2.plot)
```

Now that we know the vast majority of each promoter encountered is in fact
the claimed promoter, we can simply remove all instances where the promoter
obtained from sequencing does not match the claimed promoter.

```{r}
additional <- additional %>% filter(promoter == claimed_promoter)
```

Determine how depleted the strains became just after mating them into PA14.

```{r, fig.width=10, fig.height=5}
additional <- additional %>% 
	group_by(spacer, promoter, strain) %>% 
	summarise(count = sum(count)) %>% 
	pivot_wider(
		id_cols = "spacer", 
		names_from = c("promoter", "strain"), 
		values_from = "count", 
		values_fill = 0) %>%
	pivot_longer(
		!spacer, 
		names_to = c("cloned_strain"), 
		values_to = "count")

additional.cpm <- additional %>% 
	pivot_wider(
		id_cols = "spacer", 
		names_from = c("cloned_strain"), 
		values_from = "count") %>%
	ungroup %>%
	select(-spacer) %>% 
	data.matrix %>%
	cpm %>%
	as_tibble %>%
	bind_cols(
		additional %>% pivot_wider(
			id_cols = "spacer",
			names_from = c("cloned_strain"),
			values_from = "count") %>%
			select(spacer)) %>% 
	inner_join(
		annotated_key %>% 
			select(name, type), 
		by = c("spacer" = "name")) %>% 
	pivot_longer(
		!c(spacer, type), 
		names_to = "cloned_strain", 
		values_to = "CPM") %>% 
	mutate(`Cloned Strain` = factor(cloned_strain)) %>% 
	mutate(`Type` = factor(type)) %>% 
	group_by(`Cloned Strain`) %>% 
	nest

cpm.max <- additional.cpm %>% unnest(data) %>% ungroup %>% select(CPM) %>% max

additional.cpm <- additional.cpm %>%
	mutate(
		plot = map2(
			data,
			`Cloned Strain`,
			~ggplot(
				data = .x,
				aes(x = CPM, fill = Type)) +
				geom_density() +
				theme_ipsum() +
				scale_fill_viridis(discrete = TRUE, alpha = 0.5, option = "cividis", direction = -1) +
				scale_x_continuous(
					trans = scales::pseudo_log_trans(base = 10),
					breaks = c(0, 10^(1:6)),
					labels = label_number_si(),
					limits = c(0, cpm.max)) +
				theme(legend.position = "bottom") +
				ggtitle(paste("Strain:", .y))))

print(additional.cpm$plot)
```